Effects of epidemiological structure on the transient evolution of HIV virulence

### 2017-01-11 ms/cover_letter/plosHIVcover_resubm2.tex
still working ...

\usepackage{graphicx}
\usepackage{myltr2}
\usepackage{times}
\usepackage[utf8]{inputenc}
%\usepackage[sort&compress]{natbib}

@@

\newcommand{\revcomment}[1]{\emph{#1}}
\newcommand{\response}[1]{#1}
\newcommand{\rzero}{\ensuremath{\mathcal R}_0}
\newcommand{\todo}[1]{{\color \red \bf #1}}

@@

@@

@@

@@

@@

@@

\response{(confusion between transmission rate and  ${\mathcal R}_0$ ---
  clarify in text)}

@@

@@

>>>>

\usepackage{graphicx}
\usepackage{myltr2}
\usepackage{times}
\usepackage{color}
\usepackage[utf8]{inputenc}
%\usepackage[sort&compress]{natbib}

@@

\newcommand{\revcomment}[1]{\emph{#1}}
\newcommand{\response}[1]{#1}
\newcommand{\rzero}{\ensuremath{\mathcal R}_0}
\newcommand{\todo}[1]{{\color{red} \bf #1}}

@@

\response{
  Because the total evolutionary change in SPVL over the course of the growth phase is
equal to the time until the epidemic peak (years) times the 
speed of virulence evolution (change in SPVL/year); the faster
evolution proceeds, the closer SPVL can get to its growth-phase optimum (i.e., to the
SPVL that maximizes $r$).
}

@@

\response{
\todo{diagram of models}
}

@@

\response{
For clarity, we have changed Figure 5 to show the scaled parameters.
}

@@

\response{

}

@@

\response{
Thank you!
}

@@

\response{(confusion between transmission rate and  ${\mathcal R}_0$ ---
  clarify in text)
}

@@

\response{
  We have changed this wording to ``contact structure'' throughout
}

@@

\response{
  added \ldots
}

### 2017-01-11 ms/ms.tex
still working ...

\newcommand{\todo}[1]{\textbf{#1}}

@@

Choosing the initial conditions for the simulations is something of a
challenge. In the most common theoretical modeling contexts
researchers are primarily interested in equilibria (or other long-term
dynamical attractors such as limit cycles) and are exploring models
that have a single stable attractor, so initial conditions can be
ignored as long as we run models for long enough.  In eco-evolutionary
dynamics, however, the initial conditions do affect our
conclusions. We have no empirical information that would justify a
particular quantitative choice of the fraction infected and the mean
and variance of the distribution of SPVL at the point when the
pandemic strain of HIV-1 entered the human population; in any case,
the level of realism of our model does not support such a detailed
consideration of the early dynamics of HIV.  In most cases, we started
with an initial \Lspvl\ of 3.0, to match the value used by
Shirreff \etal. Shirreff \etal\ use an initial $I(0)=10^{-3}$;
we show results for this value in \figurename~\ref{fig:panel3}.
Because we calibrated parameters based on the initial epidemic growth rate
(see next section), we set $I(0)$

lead to very low equilibrium prevalence, we switched to 
$I(0)=10^{-4}$ for the remainder of our analysis.
We ran one small set of sensitivity analyses with baseline values of the
epidemiological parameters (Table~1) to test the effects of the
initial conditions. Appendix~S1 provides
further details on initial conditions, such
as the initial distribution of SPVL around the mean and the
distribution of initial infected density across single people
and different partnership types.

@@

Across the entire range of parameters covered by the LHS analysis, all of the classes of models we considered produce qualitatively similar virulence trajectories, which we quantify in terms of the expected time of progression
to AIDS (\figurename~\ref{fig:virtraj}: lower progression time corresponds
to higher virulence). Although the speed of virulence evolution varies, leading to wide variation in the minimum expected progression time (means ranging from approximately 6 to 12 years), virulence peaks in all models between 200 and 300 years.

>>>>

\usepackage{color}
\newcommand{\todo}[1]{\color{red} \textbf{#1}}

@@

Choosing the initial conditions for the simulations is something of a
challenge. In the most common theoretical modeling contexts
researchers are primarily interested in equilibria (or other long-term
dynamical attractors such as limit cycles) and are exploring models
that have a single stable attractor, so initial conditions can be
ignored as long as we run models for long enough.  In eco-evolutionary
dynamics, however, the initial conditions do affect our
conclusions. We have no empirical information that would justify a
particular quantitative choice of the fraction infected and the mean
and variance of the distribution of SPVL at the point when the
pandemic strain of HIV-1 entered the human population; in any case,
the level of realism of our model does not support such a detailed
consideration of the early dynamics of HIV.  In most cases, we started
with an initial \Lspvl\ of 3.0, to match the value used by
Shirreff \etal. Shirreff \etal\ use an initial $I(0)=10^{-3}$;
we show results for this value in \figurename~\ref{fig:panel3}.
Because we calibrated parameters based on the initial epidemic growth rate
(see ``Simulation runs'' below), we set $I(0)$ to $10^{-4}$ to ensure
that the exponential growth phase lasted long enough for reliable
estimation of the initial growth rate.

We ran one small set of sensitivity analyses with baseline values of the
epidemiological parameters (Table~1) to test the effects of the
initial conditions. Appendix~S1 provides
further details on initial conditions, such
as the initial distribution of SPVL around the mean and the
distribution of initial infected density across single people
and different partnership types.

@@

Across the entire range of parameters covered by the LHS analysis, all of the classes of models we considered produce qualitatively similar virulence trajectories, which we quantify in terms of the expected time of progression
to AIDS (\figurename~\ref{fig:virtraj}: lower progression time corresponds
to higher virulence). Although the speed of virulence evolution varies, leading to wide variation in the peak SPVL (ranging from \todo{XX} to \todo{YY})), virulence peaks in all models between 200 and 300 years.
